---
title: "Classification"
author: "Jeremy Binagia and Sai Gourisankar"
date: "7/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Notes/Assumptions
- study E only I believe? 
- predict 18th week total PANSS score 
- i.e. create a sv file that contains the PatientID and the predicted 18th-week PANSS score
- can submit predictions here https://www.kaggle.com/c/stanford-stats-202-forecasting/data

# Thoughts/Conclusions

# Setup
```{r Load Libraries, results="hide"}
rm(list = ls()) # clear global environment 
library(dplyr)
```

```{r Load Data}
A_df=read.csv("Data/Study_A.csv")
B_df=read.csv("Data/Study_B.csv")
C_df=read.csv("Data/Study_C.csv")
D_df=read.csv("Data/Study_D.csv")
E_df=read.csv("Data/Study_E.csv")
summary(E_df)
length(unique(E_df$PatientID))
```
Note that most patients from study E are from USA or Russia (about an even split). 

The following creates a list of the patients we should consider for the Kaggle submission. 
```{r prediction-patients}
sample_submission_df = read.csv("Data/sample_submission_PANSS.csv")
prediction.patients = sample_submission_df$PatientID # the PatientID #s we should use for Kaggle submission 
length(unique(prediction.patients)) # 379 distinct values 
#n_distinct(prediction.patients)   # gives same result 
```

