---
title: "Treatment Effect"
author: "Jeremy Binagia and Sai Gourisankar"
date: "7/5/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls()) # clear global environment 
```

## Setup 
```{r Load Libraries}
#library(dplyr)
library(ggplot2)
library(hexbin)
library(RColorBrewer)
library(ggrepel)
library(extrafont)
library(ggfortify)
library(ggsci)
#library(forecast)
#library(zoo)
library(extrafont)
#font_import()
#loadfonts(device = "win")
```

The first thing to explore is looking at the effect of the drug treatment on the total PANSS score over time.
```{r Load Data}
A_df=read.csv("Data/Study_A.csv")
B_df=read.csv("Data/Study_B.csv")
C_df=read.csv("Data/Study_C.csv")
D_df=read.csv("Data/Study_D.csv")
E_df=read.csv("Data/Study_E.csv")

summary(E_df)
```

Note that study E does not contain the `LeadStatus` column so we remove it from the other studies before combining the various dataframes. 
```{r combine datasets}
#remove LeadStatus column
A_sub = A_df[ , -which(names(A_df) %in% c("LeadStatus"))]
B_sub = B_df[ , -which(names(B_df) %in% c("LeadStatus"))]
C_sub = C_df[ , -which(names(C_df) %in% c("LeadStatus"))]
D_sub = D_df[ , -which(names(D_df) %in% c("LeadStatus"))]
E_sub = E_df # study E does not have the LeadStatus column!

combined_df = rbind(A_sub,B_sub,C_sub,D_sub,E_sub) # combine datasets
combined_df$totalP=rowSums(combined_df[,c("P1","P2","P3","P4","P5","P6","P7")])
combined_df$totalN=rowSums(combined_df[,c("N1","N2","N3","N4","N5","N6","N7")])
combined_df$totalG=rowSums(combined_df[,c("G1","G2","G3","G4","G5","G6","G7","G8","G9","G10","G11","G12","G13","G14","G15","G16")])
```

## Visual inspection of treatment vs. control 
### Total PANSS scores 
We will first plot `PANSS_Total` vs `VisitDay`, for the `Treatment` and `Control` groups. 
```{r PANSS_Total vs Visit}
combined_df_treated=combined_df[which(combined_df$TxGroup == "Treatment"),]
combined_df_control=combined_df[which(combined_df$TxGroup == "Control"),]

# considered looking at a moving average
# moving_average_day=30
# combined_df_treated$roll=rollmean(combined_df_treated$PANSS_Total,moving_average_day,na.pad=TRUE)
# combined_df_control$roll=rollmean(combined_df_control$PANSS_Total,moving_average_day,na.pad=TRUE)

p = ggplot(NULL,aes())+
  geom_jitter(data=combined_df_treated,aes(x=VisitDay,y=PANSS_Total,col="Treatment"),size=0.5)+
  geom_jitter(data=combined_df_control,aes(x=VisitDay,y=PANSS_Total,col="Control"),size=0.5)+
  scale_color_manual(values=rev(pal_aaas("default")(2)))+
  #geom_line(data=combined_df_treated,aes(x=VisitDay,y=roll,col="Treatment"))+
  #geom_line(data=combined_df_control,aes(x=VisitDay,y=roll,col="Control"))+
  theme_minimal()+
  theme(legend.position = "top",legend.title=element_blank(),plot.title=element_text(hjust=0.5,size=10,family="Lato"),plot.subtitle=element_text(hjust=0.5,size=8,family="Lato"),text=element_text(size=10,family="Lato"))
# ggsave("Figures/panss_total_vs_visit_day",p,device = "pdf",width=6,height=6,units="in",dpi="retina")
ggsave("Figures/panss_total_vs_visit_day.png",width=6,height=4,units="in",device="png",dpi="retina")
```


```{r Plot, echo=FALSE}
print(p)
```

A plot of the total PANSS score over time does not seem to suggest a treatment effect. But to quantify this we can do a linear regression on the total PANSS score vs time, where we regress PANSS_Total onto VisitDay, and a dummy variable for Control or Treatment interacting with VisitDay: $y=\beta_0 + \beta_1*VisitDay + \beta_2*VisitDay*Treatment$

### Individual PANSS scores
Instead of looking at combined scores scores, we can look at individual positive and negative scores to see if any visually appear to show a trend with respect to treatment group. 
```{r individual PANSS vs Visit}
#Plot given PANSS score vs Visit Day, for patients and control
combined_df_treated=combined_df[which(combined_df$TxGroup == "Treatment"),]
combined_df_control=combined_df[which(combined_df$TxGroup == "Control"),]

i = 9 # to prevent an overabundance of output in the html/pdf document 
#for (i in 9:38) {
p = ggplot(NULL,aes())+
  geom_jitter(data=combined_df_treated,aes(x=VisitDay,y=combined_df_treated[,i],col="Treatment"),size=0.5)+
  geom_jitter(data=combined_df_control,aes(x=VisitDay,y=combined_df_control[,i],col="Control"),size=0.5)+
  scale_color_manual(values=rev(pal_aaas("default")(2)))+
  theme_minimal()+
  theme(legend.title=element_blank(),plot.title=element_text(hjust=0.5,size=10,family="Lato"),plot.subtitle=element_text(hjust=0.5,size=8,family="Lato"),text=element_text(size=10,family="Lato"))
print(p)
#}
```
Again do not see an appreciable difference between control and treatment. 

### Total scores for each category 
```{r total scores vs Visit}
#Plot total for each category vs Visit Day, for patients and control
combined_df_treated=combined_df[which(combined_df$TxGroup == "Treatment"),]
combined_df_control=combined_df[which(combined_df$TxGroup == "Control"),]

p = ggplot(NULL,aes())+
  geom_jitter(data=combined_df_treated,aes(x=VisitDay,y=totalP,col="Treatment"),size=0.5)+
  geom_jitter(data=combined_df_control,aes(x=VisitDay,y=totalP,col="Control"),size=0.5)+
  scale_color_manual(values=rev(pal_aaas("default")(2)))+
  theme_minimal()+
  theme(legend.title=element_blank(),plot.title=element_text(hjust=0.5,size=10,family="Lato"),plot.subtitle=element_text(hjust=0.5,size=8,family="Lato"),text=element_text(size=10,family="Lato"))
print(p)
```

### Total PANSS for a given study 
```{r total scores for given study}
study = E_sub

treated = study[which(study$TxGroup == "Treatment"),]
control = study[which(study$TxGroup == "Control"),]

p = ggplot(NULL,aes())+
  geom_jitter(data=treated,aes(x=VisitDay,y=PANSS_Total,col="Treatment"),size=0.5)+
  geom_jitter(data=control,aes(x=VisitDay,y=PANSS_Total,col="Control"),size=0.5)+
  scale_color_manual(values=rev(pal_aaas("default")(2)))+
  theme_minimal()+
  theme(legend.title=element_blank(),plot.title=element_text(hjust=0.5,size=10,family="Lato"),plot.subtitle=element_text(hjust=0.5,size=8,family="Lato"),text=element_text(size=10,family="Lato"))
print(p)
```
No individual study appears to show an effect of treatment. 

## Regression of total PANSS
```{r linear regression}
#attach(combined_df)
lm.fit=lm(PANSS_Total~VisitDay+VisitDay:TxGroup, data=combined_df)
contrasts(combined_df$TxGroup) # display encoding for treatment 
```
The dummy encoding is 1 for treatment, 0 for Control.

```{r examine linear regression}
summary(lm.fit)
```

From the linear regression, we cannot rule out the hypothesis that the treatment group (Control or Treated) had no effect on the total PANSS score over time. 

<!-- ### Exploring individual patients -->
<!-- But maybe what we need to do is look at each patient individually - given they may have started at different total PANSS (so different severities of the disease), we need to look at their individual differences in total PANSS over time, in treatment and control groups. *bookmark* -->
<!-- ```{r create delta PANSS} -->
<!-- combined_df_zero = combined_df[which(combined_df$VisitDay==0),] # 0 day values for all patients -->
<!-- combined_df$deltaPANSS = combined_df$PANSS_Total -->

<!-- for (id in combined_df_zero$PatientID){ -->

<!--   zeroValue = combined_df_zero[which(combined_df_zero$PatientID==id),]$PANSS_Total # integer zero value -->
<!--   combined_df[which(combined_df$PatientID==id),]$deltaPANSS =     combined_df[which(combined_df$PatientID==id),]$PANSS_Total - zeroValue # calculate difference from start to finish -->

<!-- } -->
<!-- ``` -->

<!-- ```{r plot delta PANSS} -->
<!-- combined_df_treated=combined_df[which(combined_df$TxGroup == "Treatment"),] -->
<!-- combined_df_control=combined_df[which(combined_df$TxGroup == "Control"),] -->

<!-- # moving_average_day=30 -->
<!-- # combined_df_treated$roll=rollmean(combined_df_treated$deltaPANSS,moving_average_day,na.pad=TRUE) -->
<!-- # combined_df_control$roll=rollmean(combined_df_control$deltaPANSS,moving_average_day,na.pad=TRUE) -->

<!-- p=ggplot(NULL,aes())+ -->
<!--   geom_jitter(data=combined_df_treated,aes(x=VisitDay,y=deltaPANSS,col="Treatment"),size=0.5)+ -->
<!--   geom_jitter(data=combined_df_control,aes(x=VisitDay,y=deltaPANSS,col="Control"),size=0.5)+ -->
<!--   scale_color_manual(values=rev(pal_aaas("default")(2)))+ -->
<!--   #geom_line(data=combined_df_treated,aes(x=VisitDay,y=roll,col="Treatment"))+ -->
<!--   #geom_line(data=combined_df_control,aes(x=VisitDay,y=roll,col="Control"))+ -->
<!--   theme_minimal()+ -->
<!--   theme(legend.title=element_blank(),plot.title=element_text(hjust=0.5,size=10,family="Lato"),plot.subtitle=element_text(hjust=0.5,size=8,family="Lato"),text=element_text(size=10,family="Lato")) -->
<!-- print(p) -->
<!-- ``` -->
<!-- All this did was rescale the data. But we can re-do the linear regression and see if there is an effect of treatment on total PANSS score difference for each patient: -->
<!-- ```{r delta linear regression} -->
<!-- #attach(combined_df) -->
<!-- lm.fit=lm(deltaPANSS~VisitDay+VisitDay:TxGroup) -->
<!-- summary(lm.fit) -->
<!-- ``` -->

### Regression for total category scores
We still cannot rule out that there is no effect of treatment on the patient group's change in total PANSS score. Perhaps we can compare just the total positive symptoms of the patient, the total negative symptoms of the patient, and the total general psychopathology symptoms of the patient, to see if there is an effect on those scores. 
```{r individual scores}
combined_df$totalP=rowSums(combined_df[,c("P1","P2","P3","P4","P5","P6","P7")])
combined_df$totalN=rowSums(combined_df[,c("N1","N2","N3","N4","N5","N6","N7")])
combined_df$totalG=rowSums(combined_df[,c("G1","G2","G3","G4","G5","G6","G7","G8","G9","G10","G11","G12","G13","G14","G15","G16")])
#attach(combined_df)
lm.fit.p=lm(totalP~VisitDay+VisitDay:TxGroup, data=combined_df)
lm.fit.n=lm(totalN~VisitDay+VisitDay:TxGroup, data=combined_df)
lm.fit.g=lm(totalG~VisitDay+VisitDay:TxGroup, data=combined_df)
```

```{r summary of individual lin regs}
summary(lm.fit.p)
summary(lm.fit.n)
summary(lm.fit.g)
```

Individually, it does not look like there was a statistically significant effect of treatment on the total positive, negative, or general pathophysiology scores. 

